---
title: "p8105_hw6_cml2270"
author: "Catherine Lucey"
date: "11/30/2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Setup
Only done once, the setup code loads the necessary libraries and double checks that the working directory is correct. It also sets defaults for figure size with knitr, as well as a default theme and color scale for all ggplots.

```{r setup}

library(tidyverse)
library(modelr)

getwd()

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1: Birthweight Data

```{r load_clean, error = TRUE}

birthweight_df =
  read_csv("birthweight.csv") %>% 
  drop_na() %>% 
  mutate(
    babysex = as.factor(babysex),
    malform = as.factor(malform),
    frace = as.factor(frace),
    frace = recode_factor(frace,
                      "1" = "White",
                      "2" = "Black",
                      "3" = "Asian",
                      "4" = "Puerto Rican",
                      "8" = "Other"),
    mrace = as.factor(mrace),
    mrace = recode_factor(mrace,
                      "1" = "White",
                      "2" = "Black",
                      "3" = "Asian",
                      "4" = "Puerto Rican",
                      "8" = "Other")
  ) %>% 
  relocate(bwt)

```

#### EDA To Generate Hypotheses for a Model

Note that only three subjects had previously had live births, so for most subjects the `parity` variable is = 0. Similarly, no subject has a recorded previous low birthweight baby (`pnumlbw`), and no subjects have a recorded previous small for gestational age baby (`pnumgsa`).

```{r eda_viz}

birthweight_df %>% 
  ggplot(aes(y = bwt, x = momage)) +
  geom_point() +
  geom_smooth()

birthweight_df %>% 
  ggplot(aes(y = bwt, x = menarche)) +
  geom_point() +
  geom_smooth()

birthweight_df %>% 
  ggplot(aes(y = bwt, x = ppwt, color = ppbmi)) +
  geom_point() +
  geom_smooth()

birthweight_df %>% 
  ggplot(aes(y = bwt, x = wtgain)) +
  geom_point() +
  geom_smooth()

birthweight_df %>% 
  ggplot(aes(x = frace, y = bwt)) +
  geom_boxplot()

birthweight_df %>% 
  ggplot(aes(x = mrace, y = bwt)) +
  geom_boxplot()

# seems like theres an association with gaweeks
birthweight_df %>% 
  ggplot(aes(y = bwt, x = gaweeks)) +
  geom_point()

birthweight_df %>% 
  ggplot(aes(y = bwt, x = fincome)) +
  geom_point() +
  geom_smooth()

# surprisingly weak association between n cigarettes and bwt
birthweight_df %>% 
  ggplot(aes(y = bwt, x = smoken)) +
  geom_point() +
  geom_smooth()

# definitely going to include baby length
birthweight_df %>% 
  ggplot(aes(y = bwt, x = blength)) +
  geom_point() +
  geom_smooth()

# big head babies bigger
birthweight_df %>% 
  ggplot(aes(y = bwt, x = bhead)) +
  geom_point() +
  geom_smooth()

birthweight_df %>% 
  ggplot(aes(y = bwt, x = babysex)) +
  geom_point() +
  geom_smooth()

```

Based on my exploratory analysis of the relationship between birthweight and the different recorded variables, I will build a model in which weight at birth depends on the size of the baby's head (`bhead`), the length of the baby (`blength`), and the gestational age in weeks (`gaweeks`), as there seems to be some sort of linear relationship between birthweight and variable.


```{r build_linear_mod}

model_one = 
  lm(bwt ~ bhead + blength + gaweeks, data = birthweight_df)

summary(model_one)

model_one %>% 
  broom::tidy() %>% 
  knitr::kable()

# plot predicted / fitted values against residuals

birthweight_df %>% 
  add_predictions(model_one) %>% 
  add_residuals(model_one) %>% 
  ggplot(aes(x = resid, y = pred)) + 
  geom_point()

```

#### Defining Alternative Models

```{r alt_models}

alt_model_one =
  lm(bwt ~ blength + gaweeks, data = birthweight_df)

alt_model_two = 
  lm(bwt ~ bhead + blength + babysex + bhead*blength + blength*babysex + bhead*babysex + bhead*blength*babysex, data = birthweight_df)

```

#### Cross Validation of the Three Models


```{r validate_mods}

print(rmse(model_one, birthweight_df))
print(rmse(alt_model_one, birthweight_df))
print(rmse(alt_model_two, birthweight_df))

```

Cross validate the three models:

```{r cv}

bwt_cv_df =
  crossv_mc(birthweight_df, 1000) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

bwt_cv_df =
  bwt_cv_df %>% 
  mutate(
    model_one = map(.x = train, ~lm(bwt ~ bhead + blength + gaweeks, data = .x)),
    alt_model_one = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    alt_model_two = map(.x = train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + blength*babysex + bhead*babysex + bhead*blength*babysex, data = .x))
  ) %>% 
  mutate(
    rmse_model_one = map2_dbl(.x = model_one, .y = test, ~rmse(model = .x, data = .y)),
    rmse_alt_one = map2_dbl(.x = alt_model_one, .y = test, ~rmse(model = .x, data = .y)),
    rmse_alt_two = map2_dbl(.x = alt_model_two, .y = test, ~rmse(model = .x, data = .y))
  )

bwt_cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    rmse_model_one:rmse_alt_two,
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_boxplot()

```

Based on the root mean squared estimates of the raw models as well as the root mean squared estimates generated by cross validation, the first model, which regresses birthweight on `bhead`, `blength`, and `gaweeks`, is the best fit. However, model_one, my proposed model, and the second alternative model (which regresses on `bhead`, `blength`, and `babysex` in place of `gaweeks`) have very similar RMSE values and are likely similarly good models.

# Problem 2: Bootstrap Central Park Data

```{r load_data, message=FALSE}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

```{r bstrap_model}

boot_sample = function(df){
  
  sample_frac(df, size = 1, replace = TRUE)
}

boot_strap_df = 
  tibble(
    bstrap_number = 1:5000, 
    bstrap_sample = rerun(5000, boot_sample(weather_df))
  )

bootstrap_results =
  boot_strap_df %>% 
  mutate(
   models = map(.x = bstrap_sample, ~lm(tmax ~ tmin, data = .x))
   )
  
beta_df =
  bootstrap_results %>%
  mutate(
    results = map(models, broom::tidy)
  ) %>% 
  select(bstrap_number, results) %>% 
  unnest(results) %>%
  pivot_wider(
    names_from = term,
    values_from = c(estimate, std.error, statistic, p.value)
  ) %>% 
  janitor::clean_names() %>% 
  mutate(
    log_b0b1 = log(estimate_intercept * estimate_tmin)
  )

confidence_interval_beta = quantile(beta_df$log_b0b1, probs = c(0.025, 0.975))

rsq_df =
  bootstrap_results %>% 
  mutate(
    results = map(models, broom::glance)
  ) %>% 
  select(bstrap_number, results) %>% 
  unnest(results)

confidence_interval_rsq = quantile(rsq_df$r.squared, probs = c(0.025, 0.975))

```

```{r plot_estimates, message=FALSE}

  beta_df %>% 
  ggplot(aes(x = log_b0b1)) +
  geom_histogram()

  rsq_df %>% 
  ggplot(aes(x = r.squared)) +
  geom_histogram()

```

The plots of the distributions of the R^2 and log(B0B1) values show that, using 5000 bootstrap samples, the distributions of each parameter are approximately normal, though both distributions are slightly left skewed.

The 95% confidence interval for log(β̂ 0∗β̂1) across 5000 bootstrap samples is (1.965, 2.059). The 95% confidence interval for r2 across 5000 bootstrap samples is (0.894, 0.928)
